\ttitle

### 2016-04-16 Chapters/Introduction.tex
Expanded aetiology

%\todo{Short discussion of etiology}
\subsection{Aetiology of pancreatic cancer}
Several factors have been identified as being associated with an increased risk of pancreatic cancer.
The most important risk factor associated with increased incidence of pancreatic cancer is tobacco. 
Obesity has also been reported to be associated with increased risk of pancreatic cancer although the underlying mechanisms are unclear.
Other possible risk factors include increasing age, diet high in saturated fats, chronic pancreatitis, family history of pancreatic cancer and genetic abnormalities resulting in cancer syndromes.
It is of note that some of the important risk factors such as smoking, obesity, poor diet and increasing age may also be associated with other medical comorbidities including cardiorespiratory diseases that may have a significant impact on patient fitness.

>>>>

%\todo{Short discussion of etiology}
\subsection{Aetiology of pancreatic cancer}
Several factors have been identified as being associated with an increased risk of pancreatic cancer.
The most important risk factor associated with increased incidence of pancreatic cancer is tobacco. 
Tobacco smoking has been reported to be associated with at least a 2-fold increased risk of pancreatic cancer and this risk increases to 5-fold in patients with over 40 pack-years of smoking \parencite{raimondi_early_2007, iodice_tobacco_2008}. 
This increased risk persists for at least 10 years after cessation of smoking \parencite{iodice_tobacco_2008}.

The association between alcohol consumption and pancreatic cancer is less clear. 
While alcohol consumption on its own has not been shown to result in an increased risk of pancreatic cancer, the incidence of pancreatic cancer in heavy drinkers is greater than in the general population.
This may be due to the confounding effect of cigarette smoking in this cohort of patients \parencite{jiao_alcohol_2009, rohrmann_ethanol_2009}.

Obesity and high caloric intake have also been reported to be associated with increased risk of pancreatic cancer although the underlying mechanisms and the effects of confounding factors are poorly understood \parencite{berrington_de_gonzalez_meta-analysis_2003, larsson_body_2007, li_body_2009}.

Other possible risk factors include increasing age, diet high in saturated fats, chronic pancreatitis, family history of pancreatic cancer and genetic abnormalities resulting in cancer syndromes \parencite{raimondi_epidemiology_2009, maisonneuve_epidemiology_2010}.
It is of note that some of the important risk factors such as smoking, obesity, poor diet and increasing age may also be associated with other medical comorbidities including cardiorespiratory diseases that may have a significant impact on patient fitness.
